<!--- Notebook section for CHES and ParlGov share covered (used with include) -->
<!--- objects "id_select", "ess_check", "tbl_file_name" defined in notebooks --->

We calculate the share of matches for the "party-voted-for" (*prtv*) question. Excluded from the calculation are instances of *other*, *independent*, and *technical* (see [Party Facts codebook](https://github.com/hdigital/partyfactsdata/blob/main/codebook/codebook.md)).

```{r}
link_table_technical <-
  read_rds("data/03-party-facts-links-technical.rds")

prtv <-
  ess_check |>
  left_join(link_table_technical, by = c("prtv" = "ess_id")) |>
  select(cntry, essround, prtv, prtv_party, all_of(id_select), partyfacts_name, technical) |>
  filter(!is.na(prtv)) |>
  mutate(is_match = if_else(is.na(.data[[id_select]]), 0, 1))

prtv_match <-
  prtv |>
  filter(technical != 7 & technical != 8 & technical != 12 | is.na(technical)) |>
  summarise(
    prvt_n = n(),
    is_match = first(is_match),
    .by = c(cntry, essround, prtv, prtv_party)
  )

prtv_share <-
  prtv_match |>
  summarise(
    share_match = (sum(prvt_n * is_match) * 100 / sum(prvt_n)) |> round(1),
    .by = c(cntry, essround)
  )
```

The table summarizes the share of party matches across all countries and ESS rounds.

```{r}
tbl_out <-
  prtv_share |>
  reframe(
    enframe(
      quantile(share_match, c(0, 0.1, 0.25, 0.5, 0.75, 1)),
      "quantile", "share_match"
    )
  ) |>
  mutate(share_match = round(share_match, 1))

write_csv(tbl_out, tbl_file_name)
tbl_out
```

The share of matched parties is weighted by the number of "party-voted-for" responses and is calculated for each country in every ESS round.

The next table summarizes the country level share of party matches for ESS rounds with data set matches.

```{r}
tbl_out <-
  prtv_share |>
  summarise(
    min = min(share_match),
    median = median(share_match) |> round(1),
    max = max(share_match),
    ess_rounds = n_distinct(essround),
    .by = cntry
  ) |>
  filter(max > 0) |>
  arrange(min, median)

if (knitr::is_html_output()) {
  tbl_out |>
    reactable(searchable = TRUE, striped = TRUE)
} else {
  tbl_out
}
```
