# CHES left-right validation

ESS linking example – see also sections "Expert survey validation" and "Performance of Party Facts linking" in manuscript.

```{r}
library(conflicted)

library(tidyverse)
conflicts_prefer(dplyr::filter, .quiet = TRUE)
library(glue)
library(knitr)

library(ggforce) # violin plots
library(reactable)
```

```{r}
ches_raw <- haven::read_dta("data-raw/ches-1999-2019_CHES_dataset_meansv3-td4m.dta")
ess_raw <- read_rds("data/05-ess-ches.rds")
ess_all_raw <- read_rds("data/02-ess-select.rds")
```

## CHES information

[Chapel Hill Expert Survey](https://www.chesdata.eu/) (CHES) series

```{r}
ches_raw |>
  summarise(
    countries = n_distinct(country),
    parties = n(),
    .by = year
  )
```


## CHES and ESS

Number of countries and parties that are included in ESS and CHES for an ESS round.

```{r}
ess_raw |>
  summarize(
    year = first(year),
    countries_n = n_distinct(cntry),
    parties_n = n_distinct(prtv),
    ches_year = first(ches_year),
    ches_parties_n = n_distinct(ches_id),
    .by = c(essround)
  ) |>
  arrange(essround)
```

Overview country coverage ESS and CHES trend file

- __ess_cntry__ --- number of countries in ESS round
- __ches_cntry__ --- number of ESS round countries in CHES
- __ches_missing__ --- names of ESS round countries not in CHES

```{r}
cntry_ches_n <-
  ess_raw |>
  summarise(ches_cntry = n_distinct(cntry), .by = essround)

cntry_ches_none <-
  ess_all_raw |>
  distinct(essround, cntry) |>
  anti_join(ess_raw |> distinct(essround, cntry)) |>
  arrange(cntry) |>
  summarise(ches_missing = paste0(cntry, collapse = ", "), .by = essround)

ess_all_raw |>
  summarise(ess_cntry = n_distinct(cntry), .by = essround) |>
  left_join(cntry_ches_n) |>
  left_join(cntry_ches_none) |>
  arrange(essround)
```

## Country-year correlation

```{r}
ess_party <-
  ess_raw |>
  filter(!is.na(prtv)) |>
  summarise(
    lr_mean = mean(lrscale, na.rm = TRUE),
    lr_sd = sd(lrscale, na.rm = TRUE),
    lr_n = n(),
    ches_lr = first(ches_lr),
    .by = c(cntry, year, prtv, prtv_party, ches_year),
  )
```

Country-year correlations for ESS and CHES left-right positions in each ESS round with at least 3 parties and 10 responses per party in a country.

ESS left-right party positions are calculated as mean values of __lrscale__ variables for respondents that voted for the party (__prtv\*__).

- __lrscale__ --- self-placement on left right scale // ESS rounds
- __prtv\*__ --- "party-voted-for" in last national election // ESS rounds
- __ches_lr__ --- left-right party position // CHES trend file
  - lrgen --- "position of the party in YEAR in terms of its overall ideological stance." // CHES trendfile

```{r}
ess_party_filter <-
  ess_party |>
  na.omit() |>
  filter(lr_n >= 10) |>
  mutate(n = n(), .by = c(cntry, year)) |>
  filter(n >= 3)

corr <-
  ess_party_filter |>
  summarise(
    n_parties = first(n),
    lr_corr = cor(lr_mean, ches_lr, use = "pairwise.complete.obs"),
    .by = c(cntry, year, ches_year)
  ) |>
  arrange(cntry, year)

# fivenum(corr$lr_corr)
```

The table summarizes the country-year correlations by providing 0%, 10%, 25%, 50%, 75%, and 100% quantiles. The results are visualized in @fig-corr-violin.

```{r}
tbl_out <-
  corr |>
  summarise(
    p0 = quantile(lr_corr, 0),
    p10 = quantile(lr_corr, 0.10),
    p25 = quantile(lr_corr, 0.25),
    p50 = quantile(lr_corr, 0.50),
    p75 = quantile(lr_corr, 0.75),
    p100 = quantile(lr_corr, 1),
    .by = year
  ) |>
  mutate(across(p0:p100, \(x) round(x, 2))) |>
  rename(ess_year = year) |>
  arrange(ess_year)

write_csv(tbl_out, "figures-tables/table-3_ches-validation.csv")
tbl_out
```

::: {.content-visible when-format="html"}
Country-year correlations by country // interactive table
:::

```{r}
if (knitr::is_html_output()) {
  corr |>
    mutate(lr_corr = round(lr_corr, 2)) |>
    reactable(searchable = TRUE, striped = TRUE)
}
```

```{r}
ro_2008 <-
  ess_party |>
  filter(lr_n >= 10) |>
  filter(cntry == "RO", year == 2008) |>
  left_join(ess_raw |> distinct(prtv, prtv_party)) |>
  mutate(lr_mean = round(lr_mean, 2)) |>
  select(-lr_sd)

ro_2008_cor <-
  cor(ro_2008$lr_mean, ro_2008$ches_lr, use = "pairwise.complete.obs") |>
  round(2)
```

Lowest country-year correlation (`r ro_2008_cor`) for ESS Romania 2008.

```{r}
ro_2008 |>
  arrange(prtv) |>
  select(cntry, year, prtv, prtv_party, lr_n, lr_mean, ches_year, ches_lr)
```

```{r}
#| label: fig-corr-violin
#| fig-cap: "Violin plot for country wise correlations ( < 0.5 removed)"

corr |>
  filter(lr_corr >= 0.5) |>
  mutate(year = factor(year)) |>
  ggplot(aes(year, lr_corr)) +
  geom_violin(colour = "darkgrey") +
  geom_sina() +
  theme_minimal()
```


## Share covered

<!--- Share covered notebook section -->

```{r}
id_select <- "ches_id" # "ches_id" + "parlgov_id"
ess_check <- ess_raw # ess_raw + ess_cabinet_raw
tbl_file_name <- "figures-tables/table-2a_ches-coverage.csv"
```

{{< include _share-covered.qmd >}}
