# ESS data sources

Information on ESS data sources used – see also section "ESS party data structure" in manuscript.

```{r}
library(tidyverse)
library(knitr)
```

## ESS data sources

ESS data sets from [europeansocialsurvey.org/data](https://www.europeansocialsurvey.org/data/)

DOI references

- ESS Round 1 – <https://doi.org/10.21338/ess1e06_6>
- ESS Round 2 – <https://doi.org/10.21338/ess2e03_6>
- ESS Round 3 – <https://doi.org/10.21338/ess3e03_7>
- ESS Round 4 – <https://doi.org/10.21338/ess4e04_5>
- ESS Round 5 – <https://doi.org/10.21338/ess5e03_4>
- ESS Round 6 – <https://doi.org/10.18712/ess6e02_5>
- ESS Round 7 – <https://doi.org/10.21338/ess7e02_2>
- ESS Round 8 – <https://doi.org/10.21338/ess8e02_2>
- ESS Round 9 – <https://doi.org/10.21338/ess9e03_1>
- ESS Round 10 – <https://doi.org/10.21338/ess10e03_1>
- ESS Round 10 – <https://doi.org/10.21338/ess10sce03_0> (self-completion)

Data files are imported into R with [readstata13](https://sjewo.github.io/readstata13/)

```{r}
tibble(ESS_file = fs::dir_ls("data-raw/", glob = "*ESS*.dta")) |>
  mutate(
    Round = str_extract(ESS_file, "\\d+") |> as.integer(),
    hash = rlang::hash_file(ESS_file),
    ESS_file = str_remove(ESS_file, "data-raw/")
  ) |>
  relocate(Round, .before = ESS_file) |>
  arrange(Round) |>
  kable()
```

_Note_ --- The ESS-10 is released with two data files, one for the standard face-to-face interviews and another file for countries with self-completion mode due to the COVID-19 restrictions.

## ESS rounds

Summary of ESS rounds

- __n__ -- number of responses
- __n\_countries__ -- number of countries in ESS round
- __inw\_first__ and __inw\_last__ -- first and last interview

```{r}
ess_raw <- read_rds("data/02-ess-select.rds")

ess_raw |>
  summarise(
    n = n(),
    n_countries = n_distinct(cntry),
    inw_first = min(year(inw_date), na.rm = TRUE),
    inw_last = max(year(inw_date), na.rm = TRUE),
    .by = "essround"
  ) |>
  arrange(essround)
```


## Countries

```{r}
ess_raw |>
  summarise(
    n = n_distinct(cntry),
    countries = paste(unique(sort(cntry)), collapse = ", "),
    .by = "essround"
  ) |>
  arrange(essround)
```
