# prt* party variables

Information on ESS party IDs from “party-voted-for” (prtc\*) and “party-close-to” (prtc\*) questions – see also section "ESS party data structure" in manuscript.

```{r}
library(conflicted)

library(tidyverse)
conflicts_prefer(dplyr::filter, .quiet = TRUE)

library(knitr)
library(reactable)
```

```{r}
# ESS datasets in rds and with selected variables created in "01-ess-prt.R"
ess9_raw <- read_rds("data/ess-waves/ESS9e03_1.rds")
ess_raw <- read_rds("data/02-ess-select.rds")

# long format ESS prt* variables
prt_raw <- read_rds("data/01-ess-prt.rds")

# prtv* (single per country) and prtc* ESS variable with ESS numeric ID
prtvc_raw <- read_rds("data/01-ess-prtv-prtc.rds")
```


## prt\* variables

All ESS rounds include two types of survey questions with party information.

- __prtv\*__ --- "Party voted for in last national election, [country]?"
- __prtc\*__ --- "Which party feel closer to, [country]?"

```{r}
ess_raw |>
  summarise(
    n = n(),
    n_countries = n_distinct(cntry),
    n_prtv = n_distinct(prtv),
    n_prtc = n_distinct(prtc),
    .by = "essround"
  ) |>
  arrange(essround)
```


## ESS-9 example

We use the ESS-9 integrated file to describe the structure of the _prt\*_ variables in ESS data files.

Each ESS round uses country level variables for the __prt\*__ variables (e.g. _prtvtcat_ --- party-voted-for Austria ESS-9).

These __prt\*__ variables include the following elements:

- starting with __prt__
- indicating the type of _prt_ variable
  - __v__ --- "party-voted-for"
  - __c__ --- "party-close-to"
- two character __country__ code
- electoral tier number for Germany and Lithuania (_prtv\*_ only)


```{r}
ess9_prt <-
  ess9_raw |>
  select(cntry, starts_with(c("prtv", "prtc"))) |>
  pivot_longer(!cntry, names_to = "variable", values_to = "party") |>
  na.omit()

ess9_prt_n <-
  ess9_prt |>
  summarise(
    cntry = first(cntry),
    responses = n(),
    parties = paste(unique(party) |> as.character() |> sort(), collapse = ", "),
    .by = c(cntry, variable)
  )
```

ESS-9 __prt\*__ variables by country

::: {.content-visible when-format="pdf"}
Examples for the first four countries in ESS-9
:::

```{r}
tbl_out <-
  ess9_prt_n |>
  summarise(
    n = n(),
    variables = paste(variable |> sort(), collapse = ", "),
    .by = cntry
  )

if (knitr::is_html_output()) {
  tbl_out |>
    reactable(searchable = TRUE, striped = TRUE)
} else {
  tbl_out |>
    slice((1:5))
}
```

Germany and Lithuania include multiple _prtv\*_ variables asking for voting decisions in each electoral tier. These variables include a number for the tier in the variable name.

We use the national tier (_"prtvede2", "prtvblt1"_) as the primary "party-voted-for" variable.

```{r}
ess9_prt |>
  filter(str_detect(variable, "\\d")) |>
  distinct(cntry, variable) |>
  arrange(cntry, variable)
```


## prtv\* ID differences

ESS may use different IDs across ESS rounds

```{r ids-rounds}
prtv_ids <-
  prt_raw |>
  mutate(party = as.character(party)) |>
  filter(
    str_detect(variable, "^prtv"), # party-voted-for variable
    !str_detect(variable, "de1|lt[23]") # national vote only DEU and LTU
  ) |>
  select(-idno, -variable, -ess_id) |>
  distinct() |>
  arrange(party_id) |>
  pivot_wider(
    names_from = party_id,
    values_from = party,
    names_prefix = "id_"
  ) |>
  arrange(cntry, essround)
```

```{r}
#| eval: false
#| echo: false
reactable(prtv_ids, searchable = TRUE, striped = TRUE)
```

e.g. Netherlands _prtv\*_ rounds 1–10

```{r nld-example}
prtv_ids |>
  filter(cntry == "NL") |>
  select(essround, id_4:id_6)
```


## prtv\*/prtc\* ID differences

ESS party IDs may differ between the _prtv\*_ and _prtc\*_ variables.

```{r}
tbl <-
  prtvc_raw |>
  mutate(
    prtv_party_id = str_extract(prtv, "\\d+-\\d+") |> str_remove("\\d+-") |> as.integer(),
    prtc_party_id = str_extract(prtc, "\\d+-\\d+") |> str_remove("\\d+-") |> as.integer()
  ) |>
  filter(prtv_party_id == prtc_party_id & prtv_party != prtc_party) |>
  select(cntry, essround, party_id = prtv_party_id, prtv_party, prtc_party) |>
  distinct() |>
  arrange(essround, cntry, prtv_party)
```

```{r}
#| eval: false
#| echo: false
reactable(tbl, searchable = TRUE, striped = TRUE)
```

Examples from six countries in ESS-9

```{r}
tbl |>
  filter(essround == 9) |>
  slice(2, 5, 18, 27, 29, 38)
```

```{r}
tbl_v <-
  prtvc_raw |>
  filter(str_detect(prtv, "FI-(8|9|10)-10-v")) |>
  distinct(essround, prtv, prtv_party)

tbl_c <-
  prtvc_raw |>
  filter(str_detect(prtc, "FI-(8|9|10)-10-c")) |>
  distinct(essround, prtc, prtc_party)

tbl_out <-
  tbl_v |>
  left_join(tbl_c)

write_csv(tbl_out, "figures-tables/table-1_ess-incoherence.csv")
```
